\section{Pervasive system}
The concept of transparency of distributed systems here is extreme, even the element that makes calculation and communication, the node, in fact disappears.

Switch from mobile nodes to \textit{sensors} that have computational capabilities and are connected to Internet

\paragraph{Distributed pervasive systems}
An extension of distributed systems with the following characteristics
\begin{itemize}
    \item \textit{Unconventional nodes}, nodes that are not explicitly designed for computing but are equipped to perform computing and communication functions
    \item \textit{Adaptivity}, considers the current context
\end{itemize}
Mobile computing introduces new problems: limited resources; variance in connectivity; variable position; different types of interfaces

\paragraph{Smart device}
A device to be defined as smart must have the following characteristics: connected to Internet; run algorithms, locally or remotely, to interpret data and understand context; offer personalized services based on context

%Data acquisition
\subsection{Data acquisition}
\paragraph{Transducers}
Devices that transform one form of energy into another. Sensor captures a physical phenomenon and translates it into electricity, the actuators the opposite

\paragraph{Sensors - Input transducers}
Acquire measurements from the physical world
\begin{itemize}
    \item \textit{Physicists}, motion that measure acceleration and rotation along three axes; environmental; position sensors
    \item \textit{Virtual}, services that provide context data to remote clients
\end{itemize}

\paragraph{Actuators - Output transducers}
Make actions that affect the physical world. They are usually powered

\paragraph{Base stations}
Collect data from distributed sensing and actuating devices, may act as a gateway between networks

\paragraph{Data management}
Saving battery is the biggest problem in this context, I may need to make \textit{continuous queries}, that is, queries that continue over time to retrieve information

%Basic methods
\subsubsection{Basic methods}
\paragraph{Batch processing}
Buffering the data acquired by the sensor on the device and then sending it to the base station or remote server. No sensor-side computation but high communication costs

\paragraph{Sampling}
Do not use all sensor data

\paragraph{Overlapping sliding windows}
\textit{Time windows} are defined and real-time computation sensor side is made on these. The \textit{overlap} has the task of capturing the phenomena that occur at the edge of the time windows

%Advanced methods
\subsubsection{Advanced methods}
Improve data quality and save energy

\paragraph{In-network query processing}
Build an overlay network in the form of a reverse tree, by aggregating the data into intermediate nodes of the tree. In this way I reduce the amount of data transmitted to the base station

\paragraph{Duty-cycling}
The radio that transmits the data remains in \textit{sleep mode} for most of the time, if you have a given "abnormal" the radio \textit{wake up} and sends the data. Coordinated sleep/wakeup scheduling algorithm

\paragraph{Model-based/data-driven approaches}
The data present a strong spatio-temporal correlation. The goal is to reduce the number of samples, read or sent, while maintaining good data quality
\begin{enumerate}
    \item \textit{Data cleaning}, eliminate outliers from the data. This step is done by the base station and not by the sensor. If a data doesn’t follow this function, it’s an outlier, I don’t send it. The most likely sensor readings are compared with a statistical model, and anomalies are detected by comparing the raw sensor data with the model values.
    \item \textit{Data acquisition}, only send data that doesn’t follow that pattern, that is, that surpasses a certain $\delta$. Example: temperature
    \item \textit{Query processing}, processing queries by accessing or generating minimal amount of data
    \item \textit{Data compression}, eliminate redundancy memorizing the extremes of a range of value. Example: Poor man’s compression-MidRange, used to reduce the amount of data I have to send
    
\end{enumerate}

\newpage